\documentclass[12pt]{article}


\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{fullpage}
%\usepackage{epsfig}
%\usepackage{epstopdf}
\everymath{\displaystyle}
\usepackage{enumerate}
\usepackage{enumitem}
\usepackage[hidelinks]{hyperref}
\usepackage{xcolor}

\newif\ifans

\anstrue

\begin{document}

\begin{center}
\underline{\LARGE{Power Series}}
\end{center}

\noindent SUGGESTED REFERENCE MATERIAL:

\medskip

\noindent As you work through the problems listed below, you should reference your lecture notes and the relevant chapters in a textbook/online resource.

\medskip

\noindent EXPECTED SKILLS:

\medskip

\begin{itemize}[topsep=0pt]

\item Use sigma notation to write the Maclaurin series for a function $f(x)$.

\item Use sigma notation to write the Taylor series for a function $f(x)$ about a specified $x=x_0$.

\item Find the interval of convergence and the radius of convergence of a power series.

\item Find the domain of a function that is expressed as a power series.

\end{itemize}

\medskip

\noindent PRACTICE PROBLEMS:

\medskip

\noindent {\bf For problems 1 \& 2, use sigma notation to write the Macluarin series for the given function.}

\begin{enumerate}

\item $f(x)=\ln(1+x)$

\ifans{\fbox{$\sum_{k=1}^{\infty}{(-1)^{k+1} \text{ } \frac{x^k}{k}}$ ; Compare this to \underline{Polynomial Approximations of Functions} \#7. }} \fi

\item $f(x)=x\cos{x}$

\ifans{\fbox{$\sum_{k=0}^{\infty}{(-1)^{k} \text{ } \frac{x^{2k+1}}{(2k)!}}$}} \fi

\end{enumerate}

\noindent {\bf For problems 3 \& 4, use sigma notation to write the Taylor series for the given function about $x=x_0$.}
\begin{enumerate}
\setcounter{enumi}{2}

\item $f(x)=e^{2x}; \text{   } x_0=\ln{3}$

\ifans{\fbox{$\sum_{k=0}^{\infty}{\frac{2^k(9)}{k!} (x-\ln{3})^k}$ ; Compare this to \underline{Polynomial Approximations of Functions} \#9.  }} \fi

\item $f(x)=\sin{x}; \text{   } x_0=\frac{\pi}{2}$

\ifans{\fbox{$\sum_{k=0}^{\infty}{(-1)^{k} \text{ } \frac{\left(x-\frac{\pi}{2}\right)^{2k}}{(2k)!}}$}} \fi

\end{enumerate}

\noindent {\bf For problems 5 -- 13, find the interval of convergence and the radius of convergence $R$ for the power series.}
\begin{enumerate}
\setcounter{enumi}{4}

\item $x+x^2+x^3+x^4+\ldots$

\ifans{\fbox{$(-1, 1); R=1$.  Note that this is the power series from \underline{Infinite Series} \#24.}} \fi

\item $x-\frac{x^2}{2}+\frac{x^3}{3}-\frac{x^4}{4}+\frac{x^5}{5}-\ldots$

\ifans{\fbox{$(-1, 1]; R=1$.  This is the Maclaurin series for $f(x)=\ln(1+x)$.  See problem \#1.}} \fi

\item $\sum_{k=0}^{\infty}{\frac{x^k}{k!}}$

\ifans{\fbox{$(-\infty, +\infty); R=+\infty$.  This is the Maclaurin series for $f(x)=e^x.$}} \fi

\item $\sum_{k=0}^{\infty}{(-1)^k\frac{x^{2k+1}}{2k+1}}$

\ifans{\fbox{$[-1, 1]; R=1$.}} \fi

\item $\sum_{k=0}^{\infty}{\frac{(-5)^k x^k}{\sqrt{k+10}}}$

\ifans{\fbox{$\left(-\frac{1}{5}, \frac{1}{5}\right]; R=\frac{1}{5}$; Detailed Solution: \textcolor{blue}{\href{http://www.math.drexel.edu/classes/Calculus/resources/Math123HW/Solutions/123_12_Power_Series_09.pdf}{Here}}}} \fi

\item $\sum_{k=0}^{\infty}{\left[(2k)! \text{ } (2x+1)^k\right]}$

\ifans{\fbox{$\left[-\frac{1}{2}, -\frac{1}{2}\right]$, or just $\left\{-\frac{1}{2}\right\}; R=0$.  In other words, the series converges only when $x=-\frac{1}{2}$.   }} \fi

\item $\sum_{k=0}^{\infty}{\left[ \left( \frac{2}{7} \right)^k (x+4)^{k+1} \right] }$

\ifans{\fbox{$\left(-\frac{15}{2}, -\frac{1}{2}\right); R=\frac{7}{2}$; Detailed Solution: \textcolor{blue}{\href{http://www.math.drexel.edu/classes/Calculus/resources/Math123HW/Solutions/123_12_Power_Series_11.pdf}{Here}}}} \fi


\item $\sum_{k=0}^{\infty}{(-1)^k\frac{x^{2k+1}}{(2k+1)!}}$

\ifans{\fbox{$(-\infty, +\infty); R=+\infty$.  This is the Maclaurin series for $f(x)=\sin{x}$.}} \fi

\item $\sum_{k=2}^{\infty}{\frac{(x-3)^k}{k\ln{k}}}$

\ifans{\fbox{$[2, 4); R = 1$}} \fi

\end{enumerate}

\noindent {\bf For problems 14 -- 16, a function is represented as a power series.  Find the domain of the function.}
\begin{enumerate}
\setcounter{enumi}{13}

\item $f(x)=\sum_{k=0}^{\infty}{\left[(-1)^{k+1} \text{ } (x-2)^k\right] }$

\ifans{\fbox{\parbox{1\linewidth}{The domain of $f(x)$ is $1<x<3$.  This is the Taylor series for $f(x)=\frac{1}{1-x}$ about $x=2$.  See \underline{Polynomial Approximations of Functions} \#8.  }}} \fi

\item $f(x)=\sum_{k=0}^{\infty}{(-1)^k\frac{x^{2k}}{(2k)!} }$

\ifans{\fbox{\parbox{1\linewidth}{The domain of $f(x)$ is all real numbers.  This is the Maclaurin series for $f(x)=\cos{x}$.}}} \fi

\item $f(x)=\sum_{k=0}^{\infty}{\frac{e^{(k^2)}x^k}{k!} }$

\ifans{\fbox{The domain of $f(x)$ is only $x=0$.; Detailed Solution: \textcolor{blue}{\href{http://www.math.drexel.edu/classes/Calculus/resources/Math123HW/Solutions/123_12_Power_Series_16.pdf}{Here}}}} \fi


\end{enumerate}


\end{document}