\documentclass[12pt]{article}


\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{fullpage}
%\usepackage{epsfig}
%\usepackage{epstopdf}
\everymath{\displaystyle}
\usepackage{enumerate}
%\usepackage[hidelinks]{hyperref}
\usepackage{enumitem}
\usepackage[hidelinks]{hyperref}
\usepackage{xcolor}

\newif\ifans

\ansfalse

\begin{document}

\begin{center}
\underline{\LARGE{The Comparison, Limit Comparison, Ratio, \& Root Tests}}
\end{center}

\noindent SUGGESTED REFERENCE MATERIAL:

\medskip

\noindent As you work through the problems listed below, you should reference your lecture notes and the relevant chapters in a textbook/online resource.

\medskip

\noindent EXPECTED SKILLS:

\medskip

\begin{itemize}[topsep=0pt]

\item Use the following tests to make a conclusion about the convergence of series with no negative terms:

- Comparison Test 

- Limit Comparison Test 

- Ratio Test 

- Root Test 

\end{itemize}

\medskip

\noindent PRACTICE PROBLEMS:

\medskip

\noindent {\bf For problems 1 \& 2, apply the Comparison Test to determine if the series converges.  Clearly state to which other series you are comparing.}

\begin{enumerate}

\item $\sum_{k=1}^{\infty}{\frac{1}{3^k+5}}$

\ifans{\fbox{\parbox{1\linewidth}{$\frac{1}{3^k+5}<\frac{1}{3^k}$ for $k\geq1$.  \\ \\ Since $\sum_{k=1}^{\infty}{\frac{1}{3^k}}$ converges (geometric series, $\textstyle r=\frac{1}{3}$), $\sum_{k=1}^{\infty}{\frac{1}{3^k+5}}$ must converge.}}} \fi

\item $\sum_{k=3}^{\infty}{\frac{1}{3(k-2)}}$

\ifans{\fbox{\parbox{1\linewidth}{$\frac{1}{3(k-2)}=\frac{1}{3k-6}>\frac{1}{3k}$ for $k\geq3$.  \\ \\ Since $\sum_{k=3}^{\infty}{\frac{1}{3k}}$ diverges ($p$-series with $p=1$), $\sum_{k=3}^{\infty}{\frac{1}{3(k-2)}}$ must diverge as well.}}} \fi

\end{enumerate}

\noindent {\bf For problems 3 \& 4, apply the Limit Comparison Test to determine if the series converges.  Clearly state to which other series you are comparing.}

\begin{enumerate}
\setcounter{enumi}{2}

\item $\sum_{k=1}^{\infty}{\frac{1}{3(k+2)}}$

\ifans{\fbox{\parbox{1\linewidth}{$\lim_{k \rightarrow \infty}{\frac{\frac{1}{3(k+2)}}{\frac{1}{k}}}=\lim_{k \rightarrow \infty}{\frac{k}{3k+6}}=\frac{1}{3}$, which is finite and nonzero. 
\\ \\ So since $\sum_{k=1}^{\infty}{\frac{1}{k}}$ diverges (Harmonic Series), $\sum_{k=1}^{\infty}{\frac{1}{3(k+2)}}$ must diverge as well.   }}} \fi

\item $\sum_{k=2}^{\infty}{\frac{1}{3^k-5}}$

\ifans{\fbox{\parbox{1\linewidth}{$\lim_{k \rightarrow \infty}{\frac{\frac{1}{3^k-5}}{\frac{1}{3^k}}}=\lim_{k \rightarrow \infty}{\frac{3^k}{3^k-5}}=1$, which is finite and nonzero. 
\\ \\ So since $\sum_{k=2}^{\infty}{\frac{1}{3^k}}$ converges (geometric series, $\textstyle r=\frac{1}{3}$), $\sum_{k=2}^{\infty}{\frac{1}{3^k-5}}$ must converge.   }}} \fi


\end{enumerate}

\noindent {\bf For problems 5 -- 7, apply the Ratio Test to determine if the series converges.  If the Ratio Test is inconclusive, apply a different test.}

\begin{enumerate}
\setcounter{enumi}{4}

\item $\sum_{k=0}^{\infty}{\frac{1}{k!}}$

\ifans{\fbox{The series converges by the Ratio Test.}} \fi

\item $\sum_{k=0}^{\infty}{\frac{1}{3^k}}$

\ifans{\fbox{\parbox{1\linewidth}{The series converges by the Ratio Test. [Of course this just confirms what we already knew as this is a geometric series with $\textstyle r=\frac{1}{3}$.] }}} \fi

\item $\sum_{k=1}^{\infty}{\frac{1}{3(k+2)}}$

\ifans{\fbox{\parbox{1\linewidth}{The Ratio Test is inconclusive; however, as shown in \#3 above, the series diverges by the Limit Comparison Test.}}} \fi

\end{enumerate}

\noindent {\bf For problems 8 -- 10, apply the Root Test to determine if the series converges.  If the Root Test is inconclusive, apply a different test.}

\begin{enumerate}
\setcounter{enumi}{7}

\item $\sum_{k=0}^{\infty}{\frac{1}{3^k}}$

\ifans{\fbox{\parbox{1\linewidth}{The series converges by the Root Test. [Of course this just confirms what we already knew as this is a geometric series with $\textstyle r=\frac{1}{3}$.] }}} \fi

\item $\sum_{k=1}^{\infty}{\left(1+\frac{1}{k}\right)^k}$

\ifans{\fbox{\parbox{1\linewidth}{The Root Test is inconclusive; however, as shown in the previous assigned problems, \underline{Convergence Tests} \#7, the series diverges by the Divergence Test.}}} \fi

\item $\sum_{k=1}^{\infty}{\frac{k}{7^k}}$

\ifans{\fbox{The series converges by the Root Test.; Detailed Solution: \textcolor{blue}{\href{http://www.math.drexel.edu/classes/Calculus/resources/Math123HW/Solutions/123_09_Comparison_Ratio_Root_Tests_10.pdf}{Here}}}} \fi 

\end{enumerate}

\noindent {\bf For problems 11 -- 22, apply the Comparison Test, Limit Comparison Test, Ratio Test, or Root Test to determine if the series converges.  State which test you are using, and if you use a comparison test, state to which other series you are comparing to.}

\begin{enumerate}
\setcounter{enumi}{10}

\item $\sum_{k=1001}^{\infty}{\frac{1}{\sqrt[3]{k}-10 }}$

\ifans{\fbox{The series diverges by the Comparison Test.  Compared to $\sum_{k=1001}^{\infty}{\frac{1}{\sqrt[3]{k}}}$.}} \fi

\item $\sum_{k=1}^{\infty}{\frac{4k^2+5k}{\sqrt{10+k^5} }}$

\ifans{\fbox{\parbox{1\linewidth}{The series diverges by the Limit Comparison Test.  Compared to $\sum_{k=1}^{\infty}{\frac{1}{\sqrt{k}}}$.; \\Detailed Solution: \textcolor{blue}{\href{http://www.math.drexel.edu/classes/Calculus/resources/Math123HW/Solutions/123_09_Comparison_Ratio_Root_Tests_12.pdf}{Here}}}}} \fi



\item $\sum_{k=0}^{\infty}{\frac{2k+1}{(2k)! }}$

\ifans{\fbox{The series converges by the Ratio Test.}} \fi

\item $\sum_{k=1}^{\infty}{\frac{k^2\cos^2k}{2+k^5 }}$

\ifans{\fbox{The series converges by the Comparison Test.  Compared to $\sum_{k=1}^{\infty}{\frac{1}{k^3}}$.}} \fi

\item $\sum_{k=1}^{\infty}{\frac{1}{k^{(5k)}}}$

\ifans{\fbox{The series converges by the Root Test.}} \fi

\item $\sum_{k=0}^{\infty}{\frac{2+2^k}{3+3^k }}$

\ifans{\fbox{The series converges by the Limit Comparison Test.  Compared to $\sum_{k=0}^{\infty}{\left(\frac{2}{3}\right)^k     }$.}} \fi

\item $\sum_{k=0}^{\infty}{\frac{6^{(k+1)}}{k! }}$

\ifans{\fbox{The series converges by the Ratio Test.}} \fi

\item $\sum_{k=0}^{\infty}{\frac{6^k+k}{k!+6 }}$ [Hint:Use the result from the previous problem.]

\ifans{\fbox{The series converges by the Comparison Test.  Compared to $\sum_{k=0}^{\infty}{\frac{6^{(k+1)}}{k! }}$.; Detailed Solution: \textcolor{blue}{\href{http://www.math.drexel.edu/classes/Calculus/resources/Math123HW/Solutions/123_09_Comparison_Ratio_Root_Tests_18.pdf}{Here}}}} \fi

\item $\sum_{k=2}^{\infty}{\frac{k^3-2}{\left(k^2+1\right)^2 }}$

\ifans{\fbox{The series diverges by the Limit Comparison Test.  Compared to $\sum_{k=2}^{\infty}{\frac{1}{k}}$.}} \fi

\item $\sum_{k=1}^{\infty}{\frac{\arctan{k}}{k^{1.5} }}$

\ifans{\fbox{The series converges by the Comparison Test.  Compared to $\sum_{k=1}^{\infty}{\frac{\pi/2}{k^{1.5}}   }$.}} \fi

\item $\sum_{k=1}^{\infty}{\frac{7k}{k^2+|\sin{k}| }}$

\ifans{\fbox{The series diverges by the Limit Comparison Test.  Compared to $\sum_{k=1}^{\infty}{\frac{1}{k}}$.}} \fi

\item $\sum_{k=0}^{\infty}{\frac{(2k)!}{(k!)^2 }}$

\ifans{\fbox{The series diverges by the Ratio Test.}} \fi

\end{enumerate}

\end{document}